---
title: "National GDP Growth and Changes of Stock Market Indices"
subtitle: "Evidence from France, Germany, Great Britain and Japan"
author:
- Daniel Limberg
- Lukas Mueller
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract:  | 
  Using data from Germany, Japan, Great Britain and France we provide an analysis of the effect of national GDP Growth on changes of national stock market indices. Controling for various covariates we find that for Great Britain and France GDP growth has statistically significant effects on the FTSE100 and the CAC40. 
output: 
    pdf_document:
        fig_width: 7
        fig_height: 6
        fig_caption: true
bibliography:
    - Literature.bib
toc: yes
---

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
# Descriptive & Inferential Statistics
library(stargazer)
library(Rmisc)
library(ggplot2)
library(magrittr)
library(grid)
library(DAAG)

# Set Working Directory
try(setwd("/Users/Lukas/Documents/Git/FinalPaper/"),silent=TRUE)
try(setwd("C:/Users/Dani/Documents/GitHub2/FinalPaper/"),silent=TRUE)
getwd()


# Dynamical Link to first R script file
source("FinalPaper_Code/FinalPaper_Variables.R")
```

\pagebreak

# Introduction

&nbsp;&nbsp;In this assignment, we investigate on an odd relation: GDP growth and national equity prices; that is national stock indices. The intuitive assumption that one has automatically is fairly straight forward: The better the economy performs, the higher expectations of market participants, investments and trust in the future. In turn, we should - naturally - obtain an increase the level of national stock market indices. 

&nbsp;&nbsp;However, the more interesting question is beyond this intuitive assumption:
Focusing on four OECD countries (Germany, France, Great Britain and Japan) we investigate on the question whether there are different levels of statistical significance for the effect of GDP growth on equity prices. Note that we do not compare the size of the effects (i.e. coefficients) since we investigate on different dependent variables. Rather, we compare the statistical significance across the four countries of interest. For equity prices, we take the four major national stock market indices of the countries as our dependent variable: The CAC40 (France), DAX (Germany), FTSE100 (Great Britain) and the NIKKEI (Japan). We control for a row of other, lagged explanatory determinants - not only from within the national economy (e.g. unemployment rate) but also external, more global factors (e.g. oil prices). We elaborate on the specific covariates below. For our analysis we focus on the time period between the second quarter of 1999 (where ECB interest rates became relevant) and fourth quarter of 2015. 

&nbsp;&nbsp;The rest of this paper is structured as follows:

\pagebreak

# Literature Review

&nbsp;&nbsp;We would like to provide a two-step literature review. In the first part we summarize general findings of the field regarding national indices: What are determinants that ought to be influential on the level of National Indices?
In the second, more narrow step, we focus on the relationsship between GDP growth and national indices: Did previous studies find a relationship after all? Were causal assumptions confirmed?

## National Indices

&nbsp;&nbsp;Dimson et al. provide a solid and helpful introduction to the subject ^[The work of Dimson et al. was regularly updated. The 2009 edition is the most recent version available] with a focus on world's stock markets and the question of returns [@dimson2009triumph]. 
The field provides a tremendous amount of research on the question what influences stock markets in general and national indices in specific. Work includes research on herding effects in advanced [@chiang2010empirical], Chinese [@tan2008herding] or Japanese stock markets [@chang2006idiosyncratic]. 

Other scholars aim to predict the direction of indices via neural networks - as an attempt to go beyond traditional (non-)linear models. Examples of this stream of research include the work of Guresen et al. [@guresen2011using] or a study of the Istanbul stock exchange index [@kara2011predicting]. Other authors investigate the effect of the media [@tetlock2007giving] or even the weather on stock markets [@symeonidis2010does, @kang2010weather, @yoon2009weather, @goetzmann2005rain, @cao2005stock].

  Contagion and spill-over effects are also of interest for the field. Boyer et al. provide empirical evidence that stock market crises are spread globally through asset holdings of international investors. The more open (i.e. accessible) the stock markets, the higher the co-movements, they find [@boyer2006crises]. Kenourgios et al. conclude very similar. [@kenourgios2011financial]. 

Furthermore, the field moved on to insights from behavioral economics. Bollen et al. provide an interesting piece of work and find correlations between what they call the *twitter mood* and the Dow Jones Industrial Average over time [@bollen2011twitter]. The work of Zhang et al. [@zhang2011predicting] and Si et al. [@si2013exploiting] follow a similar approach and find identical results.

## GDP Development and National Indices

&nbsp;&nbsp;The field provides various studies focusing on the relationship between GDP growth (or macroeconomic determinants of GDP growth) and stock markets in general. For example, Claessens et al. "study how local stock market development and internationalization–listing, trading, and capital raising in international exchanges–are related to economic fundamentals" [@claessens2006stock]. Duca states that there is a clear co-movement between stock market prices and GDP in developed economies. Assuming causal effects he finds that "the level of economic activity in a country can potentially depend on the stock market amongst other variables" [@duca2007relationship]. On the other hand, Ritter finds that cross-country correlation of real stock returns and per capita GDP growth over 1900-2002 is negative. This is mainly traced back to the fact that increases in capital and labor inputs (i.e. two main determinants of GDP growth) in new corporations (i.e. representing innovation) do not necessarily imply higher dividends [@ritter2005economic].

&nbsp;&nbsp;For 42 emerging markets Yartey examines the effect of institutional and macroeconomic determinants on stock market developments [@yartey2008determinants]. Similar studies focus on these determinants for a wide range of countries. These include OECD countries like Belgium [@van2006stock], France [@sraer2007performance], Germany [@antonios2010stock] or New Zealand from 1990 to 2003 [@diebold2008macroeconomic]. And developing markets such as Malaysia [@rahman2009macroeconomic] or BRICS-countries like India [@singh2010causal, @agrawalla2007causality] and South Africa [@hsing2011stock] in general. 

# Research Question and Hypothesis

&nbsp;&nbsp;&nbsp;&nbsp;*Does national GDP growth lead to growing national stock market indices?*

* H~0~: There are statistically significant effects of national GDP growth on the growth rates of national stock market indices.

* H~1~: There are no statistically significant effects of national GDP growth on the growth rates of national stock market indices.


# Data and Variables of Interest 

&nbsp;&nbsp;In the following we introduce our variables of interest and briefly summarize how they fit our purpose. All variables were collected (or aggregated from our side) on a quarterly basis.  

| **Variables**       | Number | Origin of Data | Time Range               |
|---------------------|--------|----------------|--------------------------|
| Dependent Variables | 4      | Yahoo Finance  | Quarter 2 1999 - Q4 2015 |
| Covariates          | 6      | ECB, OECD      | Q2 1999 - Q4 2015        |


## Dependent Variables

&nbsp;&nbsp;We use data from Yahoo Finance to derive four dependent variables. We are interested in equity prices in the form of national stock market indices. These indices measure the combined stock values of companies or equities in a country which are top performing according to given benchmarks. We measure the dependent variables as the change to the the average closing value of a given index. In other words, we are not interested in the *level* but rather in the *variation* of the indices. 

**Dependent Variables**   | Official Title                      | Scope
 ------------------------- | -------------------------------- | -------------
 DAX (Germany) | Deutscher Aktienindex | 30 major Companies
 CAC40 (France) | Cotation Assisté en Continu 40 | 40 highest Market Caps
 FTSE100 (Gr. Britain) | Fin. Times Stock Exchange 100 Index | 100 Companies
 NIKKEI (Japan) | NIKKEI 225 | 225 Equities

## Covariates

&nbsp;&nbsp;In the following we control for various control variables. We use a lag of one quarter and another lag of two quarters for all covariates. 

&nbsp;&nbsp;Starting with country-specific covariates (i.e. variables that are unique to France, Germany, Great Britain and France), we use OECD data to derive the three following covariates. Among them is our key independent variables: GDP growth.  

| **Country-specific Covariates** | Measured                  |   |
|---------------------------------|---------------------------|---|
| GDP growth                      | % change to previous rate |   |
| Private consumption             | in % of GDP               |   |
| Consumer price index (CPI)      | in % to previous rate     |   |

&nbsp;&nbsp;From the ECB database we use the deposit facility (i.e. interest rate for banks^[We rakishly use the term "banks" for this paper. While "Financial Institutions" would be more appropriate to apply, we dare to continue with "banks" for reasons of simplicity.] making overnight deposits with the Eurosystem). Quite noticeable is the fact that, since June 2014, the rate here is negative. That is, quite counter-intuitive from a banks are penalized when making deposits.
&nbsp;&nbsp;Again using OECD data, we take the quarterly GDP growth rate of the USA as a proxy for the general development of the world economy. The economy of the USA has a profound impact on other national business cycles, and national recessions in the US have sever repercussions across the globe. This was demonstrated again after 2007 when the US subprime mortgage crisis turned into a global recession. Moreover, the West Texas Intermediate (WTI) price changes are taken as an indicator of global economic health and a crucial determinant of economic growth.

| **Covariates**       | Measured                            |
|----------------------|-------------------------------------|
| ECB deposit facility | % change to previous rate           |
| GDP growth USA       | % change to previous rate           |
| WTI                  | % change to previous rate in U.S. $ |

&nbsp;&nbsp;At the early stage of research for this paper other covariates were taken into account: the ECB's interest rate on the main refinancing operations (MRO), the unemployment rate for the four countries of interest and the brent crude oil price. A first variance inflation factor (vif) test after prelimenary regressions suggested to ignore the mentioned variables.

# Descriptive Statistics

&nbsp;&nbsp;**Figure 1** presents a first overview of the four dependent variables from the second quarter in 1999 to the last quarter in 2015. Striking is the non-linearity of the graphs and the fact that the NIKKEI - quite contrary to the other indices - rather declined/stagnated. The DAX, CAC40 and the FTSE100 share a general upward trend.  

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
### Germany
p1 <- ggplot(data = merge10, aes(x = Date, y = DAX.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1999.2, 2015.4, 10)) + ggtitle("Germany | DAX") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### Japan
p2 <- ggplot(data = merge10, aes(x = Date, y = NIK.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1999.2, 2015.4, 10)) + ggtitle("Japan | NIKKEI") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### France
p3 <- ggplot(data = merge10, aes(x = Date, y = CAC.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1999.2, 2015.4, 10)) + ggtitle("France | CAC40") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### Great Britain
p4 <- ggplot(data = merge10, aes(x = Date, y = FTSE.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1999.2, 2015.4, 10)) + ggtitle("Great Britain | FTSE100") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of Closing Value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
```

```{r, echo=FALSE, warning=FALSE, fig.cap="Dependent Variables (Change of National Stock Indices) over time"}
multiplot(p3, p1, p4, p2, cols = 2)
```

&nbsp;&nbsp;Another interesting overview is given in **Figure 2** which summarizes the developments of our key independent variables over the same time (i.e. Q2 1999 until Q4 2015). Our country-specific, key independent variables measure GDP growth (i.e. change to previous quarter) for the country in question. Note that here we also include the graph for the USA: An external factor we assume to be influential on all four national stock indices. Compared to the dependent variables, the key independent variables show a relatively more congruent development. Remarkable is the major, well visible downturn during and in the aftermath of financial crisis in all five graphs.

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
farben <- c("firebrick1", "mediumspringgreen", "deepskyblue2", "chocolate4", "cyan2")
### Germany
p5 <- ggplot(data = merge10, aes(x = Date, y = DEU.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1999.2, 2015.4, 10)) + ggtitle("Germany") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of GDP") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### Japan
p6 <- ggplot(data = merge10, aes(x = Date, y = JPN.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1999.2, 2015.4, 10)) + ggtitle("Japan") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of GDP") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### France
p7 <- ggplot(data = merge10, aes(x = Date, y = FRA.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1999.2, 2015.4, 10)) + ggtitle("France") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of GDP") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### Great Britain
p8 <- ggplot(data = merge10, aes(x = Date, y = GBR.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1999.2, 2015.4, 10)) + ggtitle("Great Britain") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of GDP") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### USA
p9 <- ggplot(data = merge10, aes(x = Date, y = USA.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1999.2, 2015.4, 10)) + ggtitle("USA") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of GDP") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
```

```{r, echo=FALSE, warning=FALSE, fig.cap="Key Independent Variables (Quarterly GDP growth to previous quarter)"}
multiplot(p7, p5, p8, p6, p9, cols = 2)
```

&nbsp;&nbsp;**Figure 3** summarizes the main characteristics for the depent and key indepent variables.

```{r, echo=FALSE, results="asis", fig.cap="Measures of Central Tendency for DVs & Key IVs"}
# N, Mean St. Deviations, Min and Max for all Variables
stargazer(merge10[c("CAC.Close.change", "DAX.Close.change", "FTSE.Close.change", "NIK.Close.change", "L.FRA.GDP", "L.DEU.GDP", "L.GBR.GDP"), "L.JPN.GDP"], 
          type = "latex",
          title = "Descriptive statistics",
          header = FALSE,
          digits=1)
```

&nbsp;&nbsp;The according Medians for the each dependent variable are as follows: 

```{r, echo=FALSE, results="asis", fig.cap="Medians of Dependent Variables"}
for (i in 65:68) {
  merge10[, i] %>% 
    median() %>%
    paste(names(merge10)[i], ., "\n") %>%
    cat()
}
```

&nbsp;&nbsp;In the next step, **Figure 4** summarizes the frequency distributions of the dependent variables. For now, we assume the variables to be normally distributed. In following research to this assignment we will consider log-transformation for the variables.

** Question: Maybe subset data so that we only use Close. change for Q21999-2015?!**

```{r, echo=FALSE, fig.cap="Distribution of Dependent Variables: Average Closing Values"}
par(mfrow=c(2,2))
hist(merge10$CAC.Close.change,
     main="France | CAC40", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(merge10$DAX.Close.change,
     main="Germany | DAX", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(merge10$FTSE.Close.change,
     main="Great Britain | FTSE100", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
hist(merge10$NIK.Close.change,
     main="Japan | Nikkei", 
     col="black", 
     breaks = 15,
     xlab = "Average Closing Values",
     ylab = "Frequency")
```

&nbsp;&nbsp;Finally, we obtain the joint distributions between our dependent and key independent variables in **Figure 5**. Remarkable are the trends for a high GDP growth (i.e. over 2%) in the case of Germany and Japan. Here, the intuitive relation between higher GDP growth and a higher level of the index is violated. 

**QUESTION: Do we start with GDP also from Q2 1999 ?! **

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE, warning=FALSE}
#FRA
p17 <- ggplot(merge10, aes(x = L.FRA.GDP, y = CAC.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("France") +
  xlab("GDP Growth") +
  ylab("CAC Growth Rate in %") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 
#GER
p15 <- ggplot(merge10, aes(x = L.DEU.GDP, y = DAX.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Germany") +
  xlab("GDP Growth") +
  ylab("DAX Growth Rate in %") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 
#GBR
p18 <- ggplot(merge10, aes(x = L.GBR.GDP, y = FTSE.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Great Britain") +
  xlab("GDP Growth") +
  ylab("FTSE Growth Rate in %") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 
#JPN
p16 <- ggplot(merge10, aes(x = L.JPN.GDP, y = NIK.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Japan") +
  xlab("GDP Growth") +
  ylab("Nikkei Growth Rate in %") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 
```

```{r, echo=FALSE, message=FALSE, include=TRUE, cache=FALSE, warning=FALSE, fig.cap="Joint Distributions of DVs and Key IVs"}
multiplot(p15, p16, p17, p18, cols = 2)
```

# Analysis

# Basic Model

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
M1 <- lm(CAC.Close.change ~ L.FRA.GDP + L.FRA.infl + L.FRA.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP, data = merge10)

M2 <- lm(DAX.Close.change ~ L.DEU.GDP + L.DEU.infl + L.DEU.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP, data = merge10)

M3 <- lm(FTSE.Close.change ~ L.GBR.GDP + L.GBR.infl + L.GBR.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP, data = merge10)

M4 <- lm(NIK.Close.change ~ L.JPN.GDP + L.JPN.infl + L.JPN.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP, data = merge10)
```

```{r, echo=FALSE, results="asis"}
stargazer(M1, M2, M3, M4,
          type="latex",
          title = "First Model",
          dep.var.labels=c("CAC","DAX", "FTSE", "Nikkei"),
          header = FALSE,
          no.space = TRUE)
```


# Model includes year dummies

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
M5 <- lm(CAC.Close.change ~ L.FRA.GDP + L.FRA.infl + L.FRA.prvconsm + L.ECB.dep.change  + L.WTI.dollar.change + L.USA.GDP + year, data = merge10)

M6 <- lm(DAX.Close.change ~ L.DEU.GDP + L.DEU.infl + L.DEU.prvconsm + L.ECB.dep.change  + L.WTI.dollar.change + L.USA.GDP + year, data = merge10)

M7 <- lm(FTSE.Close.change ~ L.GBR.GDP + L.GBR.infl + L.GBR.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP + year, data = merge10)

M8 <- lm(NIK.Close.change ~ L.JPN.GDP + L.JPN.infl + L.JPN.prvconsm + L.ECB.dep.change  + L.WTI.dollar.change + L.USA.GDP + year, data = merge10)
```

```{r, echo=FALSE, results="asis"}
stargazer(M5, M6, M7, M8,
          type="latex",
          title = "Model with Year Dummies",
          dep.var.labels=c("CAC","DAX", "FTSE", "Nikkei"),
          header = FALSE,
          no.space = TRUE)
```

# Final Model

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
M9 <- lm(CAC.Close.change ~ L.FRA.GDP + L2.FRA.GDP + L.FRA.infl + L.FRA.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10)

M10 <- lm(DAX.Close.change ~ L.DEU.GDP + L2.DEU.GDP + L.DEU.infl + L.DEU.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10)

M11 <- lm(FTSE.Close.change ~ L.GBR.GDP + L2.GBR.GDP + L.GBR.infl + L.GBR.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10)

M12 <- lm(NIK.Close.change ~ L.JPN.GDP + L2.JPN.GDP + L.JPN.infl + L.JPN.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10)
```

```{r, echo=FALSE, results="asis"}
stargazer(M9, M10, M11, M12,
          type="latex",
          title = "Final Model with Lags for one and two quarter for GDP",
          dep.var.labels=c("CAC","DAX", "FTSE", "Nikkei"),
          header = FALSE,
          no.space = TRUE)
```


```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
summary(M13 <- lm(CAC.Close.change ~ L.FRA.GDP + L2.FRA.GDP + L.FRA.infl + L2.FRA.infl + L.FRA.prvconsm + L2.FRA.prvconsm + L.ECB.dep.change + L2.ECB.dep.change + L.WTI.dollar.change + L2.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10))

summary(M14 <- lm(DAX.Close.change ~ L.DEU.GDP + L2.DEU.GDP + L.DEU.infl + L2.DEU.infl + L.DEU.prvconsm + L2.DEU.prvconsm + L.ECB.dep.change + L2.ECB.dep.change + L.WTI.dollar.change + L2.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10))

summary(M15 <- lm(FTSE.Close.change ~ L.GBR.GDP + L2.GBR.GDP + L.GBR.infl + L2.GBR.infl + L.GBR.prvconsm + L2.GBR.prvconsm + L.ECB.dep.change + L2.ECB.dep.change + L.WTI.dollar.change + L2.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10))

summary(M16 <- lm(NIK.Close.change ~ L.JPN.GDP + L2.JPN.GDP + L.JPN.infl + L2.JPN.infl + L.JPN.prvconsm + L2.JPN.prvconsm + L.ECB.dep.change + L2.ECB.dep.change + L.WTI.dollar.change + L2.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10))
```

```{r, echo=FALSE, results="asis"}
stargazer(M13, M14, M15, M16,
          type="latex",
          title = "Final Model with Lags for one and two quarter for GDP and inflation",
          dep.var.labels=c("CAC","DAX", "FTSE", "Nikkei"),
          header = FALSE,
          no.space = TRUE)
```

```{r, echo=FALSE, message = TRUE}
DAAG::vif(M13)
DAAG::vif(M14)
DAAG::vif(M15)
DAAG::vif(M16)
```

```{r, echo=FALSE, message = TRUE}
lmtest::bptest(M13, studentize = FALSE)
lmtest::bptest(M14, studentize = FALSE)
lmtest::bptest(M15, studentize = FALSE)
lmtest::bptest(M16, studentize = FALSE)
```

```{r, echo=FALSE, message = TRUE}
lmtest::bgtest(M13, order = 2, order.by = NULL, type = c("Chisq", "F"), data = list(), fill = 0)
lmtest::bgtest(M14, order = 2, order.by = NULL, type = c("Chisq", "F"), data = list(), fill = 0)
lmtest::bgtest(M15, order = 2, order.by = NULL, type = c("Chisq", "F"), data = list(), fill = 0)
lmtest::bgtest(M16, order = 2, order.by = NULL, type = c("Chisq", "F"), data = list(), fill = 0)
```

# Reference
