---
title: "National GDP Growth and Changes of Stock Market Indices"
subtitle: "Evidence from France, Germany, Great Britain and Japan"
author:
- Daniel Limberg
- Lukas Mueller
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract:  | 
  Using data from Germany, Japan, Great Britain and France we provide an analysis of the effect of national GDP Growth on changes of national stock market indices. 
  Controling for various covariates we find that for Great Britain and France GDP growth has statistically significant effects on the FTSE100 and the CAC40. 
output: pdf_document
bibliography:
    - Literature.bib
toc: yes
---

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
# Descriptive & Inferential Statistics
library(stargazer)
library(Rmisc)
library(ggplot2)
library(magrittr)
library(grid)
library(DAAG)

# Set Working Directory
try(setwd("/Users/Lukas/Documents/Git/FinalPaper/"),silent=TRUE)
try(setwd("C:/Users/Dani/Documents/GitHub2/FinalPaper/"),silent=TRUE)
getwd()


# Dynamical Link to first R script file
source("FinalPaper_Code/FinalPaper_Variables.R")
```

\pagebreak

# Introduction

  In this assignment, we investigate on an odd relation: GDP growth and national equity prices; that is national stock indices. The intuitive assumption that one has automatically is fairly straight forward: The better the economy performs, the higher expectations of market participants, investments and trust in the future. In turn, we should - naturally - obtain an increase the level of national stock market indices. 

  However, the more interesting question is beyond this intuitive assumption:
Focusing on four OECD countries (Germany, France, Great Britain and Japan) we investigate on the question whether there are different levels of statistical significance for the effect of GDP growth on equity prices. Note that we do not compare the size of the effects (i.e. coefficients) since we investigate on different dependent variables. Rather, we compare the statistical significance across the four countries of interest. For equity prices, we take the four major national stock market indices of the countries as our dependent variable: The CAC40 (France), DAX (Germany), FTSE100 (Great Britain) and the NIKKEI (Japan). We control for a row of other, lagged explanatory determinants - not only from within the national economy (e.g. unemployment rate) but also external, more global factors (e.g. oil prices). We elaborate on the specific covariates below. For our analysis we focus on the time period between the second quarter of 1999 (where ECB interest rates became relevant) and fourth quarter of 2015. 


  The rest of this paper is structured as follows:

\pagebreak

# Literature Review

  We would like to provide a two-step literature review. In the first part we summarize general findings of the field regarding national indices: What are determinants that ought to be influential on the level of National Indices?
In the second, more narrow step, we focus on the relationsship between GDP growth and national indices: Did previous studies find a relationship after all? Were causal assumptions confirmed?

## National Indices

  Dimson et al. provide a solid and helpful introduction to the subject ^[The work of Dimson et al. was regularly updated. The 2009 edition is the most recent version available] with a focus on world's stock markets and the question of returns [@dimson2009triumph]. 
The field provides a tremendous amount of research on the question what influences stock markets in general and national indices in specific. Work includes research on herding effects in advanced [@chiang2010empirical], Chinese [@tan2008herding] or Japanese stock markets [@chang2006idiosyncratic]. 

Other scholars aim to predict the direction of indices via neural networks - as an attempt to go beyond traditional (non-)linear models. Examples of this stream of research include the work of Guresen et al. [@guresen2011using] or a study of the Istanbul stock exchange index [@kara2011predicting]. Other authors investigate the effect of the media [@tetlock2007giving] or even the weather on stock markets [@symeonidis2010does, @kang2010weather, @yoon2009weather, @goetzmann2005rain, @cao2005stock].

  Contagion and spill-over effects are also of interest for the field. Boyer et al. provide empirical evidence that stock market crises are spread globally through asset holdings of international investors. The more open (i.e. accessible) the stock markets, the higher the co-movements, they find [@boyer2006crises]. Kenourgios et al. conclude very similar. [@kenourgios2011financial]. 

Furthermore, the field moved on to insights from behavioral economics. Bollen et al. provide an interesting piece of work and find correlations between what they call the *twitter mood* and the Dow Jones Industrial Average over time [@bollen2011twitter]. The work of Zhang et al. [@zhang2011predicting] and Si et al. [@si2013exploiting] follow a similar approach and find identical results.

## GDP Development and National Indices

  The field provides various studies focusing on the relationship between GDP growth (or macroeconomic determinants of GDP growth) and stock markets in general. For example, Claessens et al. "study how local stock market development and internationalization–listing, trading, and capital raising in international exchanges–are related to economic fundamentals" [@claessens2006stock]. Duca states that there is a clear co-movement between stock market prices and GDP in developed economies. Assuming causal effects he finds that "the level of economic activity in a country can potentially depend on the stock market amongst other variables" [@duca2007relationship]. On the other hand, Ritter finds that cross-country correlation of real stock returns and per capita GDP growth over 1900-2002 is negative. This is mainly traced back to the fact that increases in capital and labor inputs (i.e. two main determinants of GDP growth) in new corporations (i.e. representing innovation) do not necessarily imply higher dividends [@ritter2005economic].

  For 42 emerging markets Yartey examines the effect of institutional and macroeconomic determinants on stock market developments [@yartey2008determinants]. Similar studies focus on these determinants for a wide range of countries. These include OECD countries like Belgium [@van2006stock], France [@sraer2007performance], Germany [@antonios2010stock] or New Zealand from 1990 to 2003 [@diebold2008macroeconomic]. And developing markets such as Malaysia [@rahman2009macroeconomic] or BRICS-countries like India [@singh2010causal, @agrawalla2007causality] and South Africa [@hsing2011stock] in general. 

# Research Question and Hypothesis

*Does national GDP growth lead to growing national stock market indices?*

* H~0~: There are statistically significant effects of national GDP growth on the growth rates of national stock market indices.

* H~1~: There are no statistically significant effects of national GDP growth on the growth rates of national stock market indices.


# Data and Variables of Interest 

We investigate on four different dependent variables: The National stock indices of Germany, France, Great Britain and Japan. We obtain data from their average closing values.

**Dependent Variables**   | Official Title                      | Scope
 -------------------- | ----------------------------------- | -----------------------
 DAX (Germany)        | Deutscher Aktienindex               | 30 major Companies
 CAC (France)         | Cotation Assisté en Continu 40      | 40 highest Market Caps
 FTSE (Great Britain) | Fin. Times Stock Exchange 100 Index | 100 Companies
 NIKKEI (Japan)       | NIKKEI 225                          | 225 Equities


  The analsis of this paper is built on unbalanced panel data that lists selected national indices against macro-economic indicators for all countries of the world, and occurrences of natural disasters from as early as 1991 until as late as 2015.

  We use National Indices from Yahoo Finance: DAX (DE), NIKKEI (JP), FTSE100 (UK), CAC40 (FR), IBOV (BR), and OMRU (RU). These indices measure the combined stock values of firms in a country which are top performing according to given benchmarka. For instance, the DAX measures the 30 most valuable firms in Germany, and the IBOV includes all Brazilian firms whose negotiabilty index added is worth 80% of all other negotiability indexes added and which have been traded the previous year. 
  Secondly, the Word Bank provides us with data on stocks traded, total value (% of GDP), foreign direct investment (% of GDP), Stocks traded, turnover ratio (%), GDP per capita growth (annual %), estimated capital stock (real 2005 USD), and GDP per capita (constant 2005 USD). The World Bank gathers its data consistetly on a yearly basis.
  Thirdly, the International Disaster Database (EM-DAT) of the Centre for Research of the Epidemiology of Disasters (CRED) provides public data on natural disasters with the aim to help development and relief agencies mitigate the impact of these disasters. We create a variable that counts the occurences of natural disasters per year in a country but do not differentiate between the types of disaster.


# Basic Model

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
M1 <- lm(CAC.Close.change ~ L.FRA.GDP + L.FRA.infl + L.FRA.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP, data = merge10)

M2 <- lm(DAX.Close.change ~ L.DEU.GDP + L.DEU.infl + L.DEU.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP, data = merge10)

M3 <- lm(FTSE.Close.change ~ L.GBR.GDP + L.GBR.infl + L.GBR.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP, data = merge10)

M4 <- lm(NIK.Close.change ~ L.JPN.GDP + L.JPN.infl + L.JPN.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP, data = merge10)
```

```{r, echo=FALSE, results="asis"}
stargazer(M1, M2, M3, M4,
          type="latex",
          title = "Basic Model",
          dep.var.labels=c("CAC","DAX", "FTSE", "Nikkei"),
          header = FALSE,
          no.space = TRUE)
```


# Model includes year dummies

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
M5 <- lm(CAC.Close.change ~ L.FRA.GDP + L.FRA.infl + L.FRA.prvconsm + L.ECB.dep.change  + L.WTI.dollar.change + L.USA.GDP + year, data = merge10)

M6 <- lm(DAX.Close.change ~ L.DEU.GDP + L.DEU.infl + L.DEU.prvconsm + L.ECB.dep.change  + L.WTI.dollar.change + L.USA.GDP + year, data = merge10)

M7 <- lm(FTSE.Close.change ~ L.GBR.GDP + L.GBR.infl + L.GBR.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP + year, data = merge10)

M8 <- lm(NIK.Close.change ~ L.JPN.GDP + L.JPN.infl + L.JPN.prvconsm + L.ECB.dep.change  + L.WTI.dollar.change + L.USA.GDP + year, data = merge10)
```

```{r, echo=FALSE, results="asis"}
stargazer(M5, M6, M7, M8,
          type="latex",
          title = "Model with Year Dummies",
          dep.var.labels=c("CAC","DAX", "FTSE", "Nikkei"),
          header = FALSE,
          no.space = TRUE)
```


# Final Model

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
M9 <- lm(CAC.Close.change ~ L.FRA.GDP + L2.FRA.GDP + L.FRA.infl + L.FRA.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10)

M10 <- lm(DAX.Close.change ~ L.DEU.GDP + L2.DEU.GDP + L.DEU.infl + L.DEU.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10)

M11 <- lm(FTSE.Close.change ~ L.GBR.GDP + L2.GBR.GDP + L.GBR.infl + L.GBR.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10)

M12 <- lm(NIK.Close.change ~ L.JPN.GDP + L2.JPN.GDP + L.JPN.infl + L.JPN.prvconsm + L.ECB.dep.change + L.WTI.dollar.change + L.USA.GDP + L2.USA.GDP + year, data = merge10)
```

```{r, echo=FALSE, results="asis"}
stargazer(M9, M10, M11, M12,
          type="latex",
          title = "Final Model with Lags for one and two quarter for GDP",
          dep.var.labels=c("CAC","DAX", "FTSE", "Nikkei"),
          header = FALSE,
          no.space = TRUE)
```

```{r, echo=FALSE}
DAAG::vif(M9)
DAAG::vif(M10)
DAAG::vif(M11)
DAAG::vif(M12)
```


# Reference
