---
title: "National GDP Growth and Changes of Stock Market Indices"
subtitle: "Evidence from France, Germany, Great Britain and Japan"
author: |
        | Daniel Limberg & Lukas Müller
        | Hertie School of Governance
date: May 2016
output: 
    html_document:
      toc: true
      toc_float:
          smooth_scroll: true
          collapsed: true
      theme: journal
      highlight: kate
      fig_caption: true
bibliography: 
- Literature.bib
- RPackages.bib
---

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
# Descriptive & Inferential Statistics
library(stargazer)
library(Rmisc)
library(ggplot2)
library(magrittr)
library(grid)
library(DAAG)
library(plotly)

# Set Working Directory
try(setwd("/Users/Lukas/Documents/Git/FinalPaper/"),silent=TRUE)
try(setwd("C:/Users/Dani/Documents/GitHub2/FinalPaper/"),silent=TRUE)
getwd()


# Dynamical Link to first R script file
source("FinalPaper_Code/FinalPaper_Variables.R")
```

\pagebreak

# Introduction

&nbsp;&nbsp;&nbsp;&nbsp;Nowadays, equity prices have become the nexus between capitalism and our daily life on the globe. Especially the financial crisis and the following European sovereign debt crisis darken the beauty of this nexus, that was once proclaimed to be a good - if not the best - example of a functioning market mechanism.\newline
In this paper, we investigate an odd relation that is a part of the nexus: GDP growth and a subset of national equity prices; that is national stock indices. The intuitive assumption that one automatically has is fairly straight forward: The better the economy performs, the higher expectations of market participants, investments and trust in the future. In turn, we should - naturally - obtain an increase in the *level* of a national stock market indices given good economic performance within a country. In a way it is reasonable to argue, that national stock indices serve as a proxy of the economy. However, the more interesting question is beyond this intuitive assumption: \newline
&nbsp;&nbsp;&nbsp;&nbsp;Focusing on four OECD countries (Germany, France, Great Britain and Japan) we investigate whether GDP growth has an effect of the *changes* of the indices. If so, are there different levels of statistical significance for the effect of GDP growth across the different countries? That is, we do not compare the size of the effects (i.e. coefficients) since we investigate on different dependent variables. We take the four major national stock market indices of the countries as our dependent variable: The CAC40 (France), DAX (Germany), FTSE100 (Great Britain) and the NIKKEI (Japan). We control for a row of other, lagged explanatory determinants - not only from within the national economy (e.g. inflation rate) but also external, more global factors (e.g. oil price). We elaborate on the specific covariates below. For our analysis we focus on the time period between the beginning of 1999 (where ECB interest rates became relevant) and the end of 2015. \newline

&nbsp;&nbsp;&nbsp;&nbsp;The rest of this paper is structured as follows. In the next section 2 we provide a brief Literature Review and introduce basic findings of the field for our topic. Section 3 provides of Research Question and Hypothesis. We then introduce our Data and Variables of Interest in section 4. We continue with some basic Descriptive Statistics (section 5) before we provide our Analysis and Discussion in section 6. We end with Limitations and opportunities for Future Research (section 7) and a Conclusion (section 8). Note that we provide diagnostics and a variation of our final model in the Appendix.

# Literature Review

&nbsp;&nbsp;&nbsp;&nbsp;We would like to provide a two-step literature review. In the first part we summarize general findings of the field regarding national indices: What are determinants that ought to be influential on the level of national indices?
In the second, more narrow step, we focus on the relationsship between GDP growth and national indices: Did previous studies find a relationship after all? Were causal assumptions confirmed?

## National Indices

&nbsp;&nbsp;&nbsp;&nbsp;Dimson et al. provide a solid and helpful introduction to the subject ^[The work of Dimson et al. was regularly updated. The 2009 edition is the most recent version available.] with a focus on world's stock markets and the question of returns [@dimson2009triumph]. The field provides a tremendous amount of research on the question what influences stock markets in general and national indices in specific. Work includes research on herding effects in advanced [@chiang2010empirical], Chinese [@tan2008herding] or Japanese stock markets [@chang2006idiosyncratic]. \newline
&nbsp;&nbsp;&nbsp;&nbsp;Other scholars aim to predict the direction of indices via neural networks - as an attempt to go beyond traditional (non-)linear models. Examples of this stream of research include the work of Guresen et al. [@guresen2011using] or a study of the Istanbul stock exchange index [@kara2011predicting]. Other authors investigate the effect of the media [@tetlock2007giving] or even the weather on stock markets [@symeonidis2010does, @kang2010weather, @yoon2009weather, @goetzmann2005rain, @cao2005stock].\newline
&nbsp;&nbsp;&nbsp;&nbsp;Contagion and spill-over effects are also of interest for the field. Boyer et al. provide empirical evidence that stock market crises are spread globally through asset holdings of international investors. The more open (i.e. accessible) the stock markets, the higher the co-movements, they find [@boyer2006crises]. Kenourgios et al. conclude very similar [@kenourgios2011financial]. Furthermore, the field moved on to insights from behavioral economics. Bollen et al. provide an interesting piece of work and find correlations between what they call the *twitter mood* and the Dow Jones Industrial Average over time [@bollen2011twitter]. The work of Zhang et al. [@zhang2011predicting] and Si et al. [@si2013exploiting] follow a similar approach and find almost identical results.

## GDP Growth and National Indices

&nbsp;&nbsp;&nbsp;&nbsp;The field provides various studies which focus on the relationship between GDP growth (or macroeconomic determinants of GDP growth) and stock markets in general. For example, Claessens et al. "study how local stock market development and internationalization listing, trading, and capital raising in international exchanges are related to economic fundamentals" [@claessens2006stock]. Duca states that there is a clear co-movement between stock market prices and GDP in developed economies. Assuming causal effects he finds that "the level of economic activity in a country can potentially depend on the stock market amongst other variables" [@duca2007relationship]. On the other hand, Ritter finds that cross-country correlation of real stock returns and per capita GDP growth over 1900-2002 is negative. This is mainly traced back to the fact that increases in capital and labor inputs (i.e. two main determinants of GDP growth) in new corporations (i.e. representing innovation) do not necessarily imply higher dividends [@ritter2005economic].\newline

&nbsp;&nbsp;&nbsp;&nbsp;For 42 emerging markets Yartey examines the effect of institutional and macroeconomic determinants on stock market developments [@yartey2008determinants]. Similar studies focus on these determinants for a wide range of countries. These include OECD countries like Belgium [@van2006stock], France [@sraer2007performance], Germany [@antonios2010stock] or New Zealand from 1990 to 2003 [@diebold2008macroeconomic]. And developing markets such as Malaysia [@rahman2009macroeconomic] or BRICS-countries like India [@singh2010causal, @agrawalla2007causality] and South Africa [@hsing2011stock] in general. 

# Research Question and Hypothesis

&nbsp;&nbsp;&nbsp;&nbsp;*Do changes in GDP growth lead to changes in the country's national stock market index?*

* H~0~: There are statistically significant effects of GDP growth on the growth rates of a country's stock market index.

* H~1~: There are no statistically significant effects of GDP growth on the growth rates of a country's stock market index.

# Data and Variables of Interest 

&nbsp;&nbsp;&nbsp;&nbsp;In the following we introduce our variables of interest and briefly summarize how they fit our purpose. All variables were collected on (or aggregated to) a quarterly format.  

| **Variables**       | Number | Origin of Data | Time Range               |
|---------------------|--------|----------------|--------------------------|
| Dependent Variables | 4      | Yahoo Finance  | Quarter 1 1999 - Q4 2015 |
| Covariates          | 6      | ECB, OECD      | Quarter 1 1999 - Q4 2015 |


## Dependent Variables

&nbsp;&nbsp;&nbsp;&nbsp;We use data from Yahoo Finance to derive four dependent variables. We are interested in equity prices in the form of national stock market indices for four OECD countries: France, Germany, Great Britain and Japan. The indices measure the combined stock values of companies and/or equities in a country which are top performing according to given benchmarks. We measure the dependent variables as the change of the average closing value from the previous quarter. In other words, we are not interested in the *level* but rather in the *variation* or *changes* of the indices over time. 

**Dependent Variables**   | Official Title                      | Scope
 ------------------------- | -------------------------------- | -------------
 CAC40 (France) | Cotation Assisté en Continu 40 | 40 highest Market Caps
 DAX (Germany) | Deutscher Aktienindex | 30 major Companies
 FTSE100 (Great Britain) | Fin. Times Stock Exchange 100 Index | 100 Companies
 NIKKEI (Japan) | NIKKEI 225 | 225 Equities

## Covariates

&nbsp;&nbsp;&nbsp;&nbsp;In the following we control for various covariates. We use a lag of one quarter and another lag of two quarters for all covariates. If not indicated differently, the default for the covariates is a lag of one quarter. \newline
&nbsp;&nbsp;&nbsp;&nbsp;For the control variables we differentiate between country-specific and general covariates that apply to all countries. Starting with country-specific covariates (i.e. variables that are unique to France, Germany, Great Britain and France), we use OECD data to derive the three following covariates. Among them is our key independent variable: GDP growth. We also use private consumption and the inflation rate (measured through the CPI) within the countries as additional control variables. 

| **Country-specific Covariates** | Measured                  |
|---------------------------------|---------------------------|
| GDP growth                      | % change to previous quarter |
| Consumer price index (CPI)      | % change to previous quarter | 
| Private consumption             | % change of share of GDP to prev. quarter |

&nbsp;&nbsp;&nbsp;&nbsp;From the ECB database we use the deposit facility (i.e. interest rate for banks^[We rakishly use the term "banks" for this paper. While "Financial Institutions" would be more appropriate to apply, we dare to continue with "banks" for reasons of simplicity.] making overnight deposits with the Eurosystem). Quite noticeable is the fact that, since June 2014, the rate is negative. That is, quite counter-intuitive, banks are penalized when making deposits.\newline
&nbsp;&nbsp;&nbsp;&nbsp;Again using OECD data, we also take the quarterly GDP growth rate of the USA as a proxy for the general development of the world economy. The economy of the USA has a profound impact on other national business cycles, and national recessions in the US have sever repercussions across the globe. This was demonstrated again after 2007 when the US subprime mortgage crisis turned into a global recession. Moreover, the West Texas Intermediate (WTI) price changes are taken as an indicator of global economic health and a crucial determinant of economic growth.

| **Covariates**        | Measured                                 |
|-----------------------|------------------------------------------|
| ECB deposit facility  | % change to previous quarter             |
| GDP growth USA        | % change to previous quarter             |
| WTI                   | % change to previous quarter             |

&nbsp;&nbsp;&nbsp;&nbsp;At the early stage of research for this paper other covariates were taken into account: the ECB's interest rate on the main refinancing operations (MRO), the unemployment rate for the four countries of interest and the brent crude oil price. However, a first variance inflation factor (vif) test after prelimenary regressions suggested to ignore the mentioned variables.

# Descriptive Statistics

## Overview Dependent Variables

&nbsp;&nbsp;&nbsp;&nbsp;Before going into our analysis, we now turn to a brief description of our available and cleaned data. \newline
&nbsp;&nbsp;&nbsp;&nbsp;Figure 1 below presents a first overview of the four dependent variables: The % change of the average closing value of the indices to the previous quarter. The graphs range from the first quarter in 1999 to the last quarter in 2015. Given the financial crisis, the CAC40, DAX, FTSE100 and NIKKEI share a remarkable but unsurprising downturn in the end of 2008 and in early 2009. 

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
### Germany
p1 <- ggplot(data = merge10, aes(x = Date, y = DAX.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1999.1, 2015.4, 5)) + ggtitle("Germany | DAX") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of av. closing value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### Japan
p2 <- ggplot(data = merge10, aes(x = Date, y = NIK.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1999.1, 2015.4, 5)) + ggtitle("Japan | NIKKEI") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of av. closing value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### France
p3 <- ggplot(data = merge10, aes(x = Date, y = CAC.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1999.1, 2015.4, 5)) + ggtitle("France | CAC40") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of av. closing value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
### Great Britain
p4 <- ggplot(data = merge10, aes(x = Date, y = FTSE.Close.change)) + geom_line() + scale_x_continuous(breaks=seq(1999.1, 2015.4, 5)) + ggtitle("Great Britain | FTSE100") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of av. closing value") + theme(legend.position="none") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8))
```

```{r, echo=FALSE, results="asis", warning=FALSE, fig.cap="Dependent variables: % change of average closing values of national stock indices over time"}
ggplotly(p3)
ggplotly(p1)
ggplotly(p4)
ggplotly(p2)
```

## Overview Independent Variables

&nbsp;&nbsp;&nbsp;&nbsp;Another interesting overview is given with Figure 2 which summarizes the developments of our key independent variables over the same time (i.e. Q1 1999 until Q4 2015). Our country-specific, key independent variables measure GDP growth (i.e. change to previous quarter) for the country in question. Remarkable again is the major, well visible downturn during and in the aftermath of financial crisis in all four graphs.

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
farben <- c("firebrick1", "mediumspringgreen", "deepskyblue2", "chocolate4", "cyan2")
### Germany
p5 <- ggplot(data = merge10, aes(x = Date, y = DEU.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1999.1, 2015.4, 5)) + ggtitle("Germany") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of GDP") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### Japan
p6 <- ggplot(data = merge10, aes(x = Date, y = JPN.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1999.1, 2015.4, 5)) + ggtitle("Japan") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of GDP") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### France
p7 <- ggplot(data = merge10, aes(x = Date, y = FRA.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1999.1, 2015.4, 5)) + ggtitle("France") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of GDP") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### Great Britain
p8 <- ggplot(data = merge10, aes(x = Date, y = GBR.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1999.1, 2015.4, 5)) + ggtitle("Great Britain") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of GDP") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
### USA
p9 <- ggplot(data = merge10, aes(x = Date, y = USA.GDP)) + geom_line() + scale_x_continuous(breaks=seq(1999.1, 2015.4, 5)) + ggtitle("USA") + theme(plot.title = element_text(size=10)) + labs(x = "Quarters", y = "% change of GDP") + theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) + theme(legend.position="none")
```

```{r, echo=FALSE, warning=FALSE, fig.cap="Key independent variables: GDP growth over time"}
multiplot(p7, p5, p8, p6, cols = 2)
```

## Key Characteristics Variables of Interest

&nbsp;&nbsp;&nbsp;&nbsp;The table below summarizes the main characteristics for the depent (National stock market indices) and key indepent (GDP Growth) variables for our analyzed time period of Q1 1999 to Q4 2015. At first glance, our N=69 in all cases seems fairly small. However, note that we aggregated our original, raw data into a quarterly format. In the original dataset we obtained around 4400 observations per regression (i.e. per index). In terms of the changes of the national indices, the DAX reached the highest (+27.6%) and the NIKKEI the lowest (-31.8%) values. Japan experienced the highest (2.7%) and Germany the lowest (-4.5%) values for GDP growth in the time period.

```{r, echo=FALSE, results="asis"}
# N, Mean St. Deviations, Min and Max for all Variables
stargazer(merge10[, c("CAC.Close.change", "DAX.Close.change", "FTSE.Close.change", "NIK.Close.change", "L.FRA.GDP", "L.DEU.GDP", "L.GBR.GDP", "L.JPN.GDP")], 
          type = "latex",
          covariate.labels = c("CAC40", "DAX", "FTSE100", "NIKKEI", "France GDP growth", "Germany GDP growth", "Great Britain GDP growth", "Japan GDP growth"),
          title = "Descriptive statistics",
          header = FALSE,
          digits=1)
```

## Frequency Distribution Dependent Variables

&nbsp;&nbsp;&nbsp;&nbsp;In the next step, the histograms below summarize the frequency distributions of the dependent variables. Please note the different scales of the y-axes and the x-axis in case of Japan. It becomes visible that all four national indices with their % change of the average closing value cluster around 0, although the CAC40 and the DAX are slightly scewed to the right. For now, we assume the variables to be normally distributed. In following research to this assignment we should, however, consider log-transformation especially in the case of Japan.
\newline

```{r, echo=FALSE, fig.cap="Dependent variables: Frequency distributions"}
par(mfrow=c(2,2))
hist(merge10$CAC.Close.change,
     main="France | CAC40", 
     col="black", 
     breaks = 15,
     xlab = "% change of average closing values",
     ylab = "Frequency")
hist(merge10$DAX.Close.change,
     main="Germany | DAX", 
     col="black", 
     breaks = 15,
     xlab = "% change of average closing values",
     ylab = "Frequency")
hist(merge10$FTSE.Close.change,
     main="Great Britain | FTSE100", 
     col="black", 
     breaks = 15,
     xlab = "% change of average closing values",
     ylab = "Frequency")
hist(merge10$NIK.Close.change,
     main="Japan | Nikkei", 
     col="black", 
     breaks = 15,
     xlab = "% change of average closing values",
     ylab = "Frequency")
```

### Joint Distributions

&nbsp;&nbsp;&nbsp;&nbsp;Below, we obtain the joint distributions between our dependent and key independent variables for the four countries. Here, we use GDP growth lagged by one quarter, which we also use in the following analysis section below. The key dependent variables on the y-axes are the % changes of average closing values for the given indices.\newline
Remarkable are the few outliers in negative GDP growth for Germany, France, and Japan, which suggest a downward trend. However, the confindence intervals already indicate that one needs to be careful with drawing conclusions from this section of the graphics. More interesting is the general upward trend for Germany, Japan, and Great Britain that starts around and above a GDP growth of 0%. This supports the original idea we described earlier on: A well performing national economy should, in theory, lead to a rise in the national stock market index. Quite to the contrary, we see rather a stagnation instead of a rise around and above a GDP growth of 0% for Japan. While we need regression modeling to draw true conclusions from this correlation, it should also be mentioned again that the NIKKEI combines the highest number of equities of all dependent variables. It is probably fair to assume that the high number of equities balances the relationship. 

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE, warning=FALSE}
#FRA
p17 <- ggplot(merge10, aes(x = L.FRA.GDP, y = CAC.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("France") +
  xlab("Lagged GDP Growth") +
  ylab("CAC40 % change of av. closing value") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 
#GER
p15 <- ggplot(merge10, aes(x = L.DEU.GDP, y = DAX.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Germany") +
  xlab("Lagged GDP Growth") +
  ylab("DAX % change of av. closing value") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 
#GBR
p18 <- ggplot(merge10, aes(x = L.GBR.GDP, y = FTSE.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Great Britain") +
  xlab("Lagged GDP Growth") +
  ylab("FTSE100 % change of average closing value") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 
#JPN
p16 <- ggplot(merge10, aes(x = L.JPN.GDP, y = NIK.Close.change)) + 
  geom_point() + 
  geom_smooth() +
  ggtitle("Japan") +
  xlab("Lagged GDP Growth") +
  ylab("NIKKEI % change of av. closing value") +
  theme(axis.text=element_text(size=8),
        axis.title=element_text(size=8)) +
  theme(plot.title = element_text(size=10)) 
```

```{r, echo=FALSE, message=FALSE, include=TRUE, cache=FALSE, warning=FALSE, fig.cap="Joint distributions of dependent and key independent variables"}
ggplotly(p17)
ggplotly(p15)
ggplotly(p18)
ggplotly(p16)
```


# Analysis and Discussion of Results

&nbsp;&nbsp;&nbsp;&nbsp;For the analysis section we provide our comprehensive model below. However, before going into the most promising model that we found during our research, we would like to offer a very basic model as a starting point of the analysis. We only regress the change of the four national stock indices on the lagged quaterly GDP growth of the respective country.

```{r, echo=FALSE, message=FALSE, include=FALSE, cache=FALSE}
M1 <- lm(CAC.Close.change ~ L.FRA.GDP, data = merge10)

M2 <- lm(DAX.Close.change ~ L.DEU.GDP, data = merge10)

M3 <- lm(FTSE.Close.change ~ L.GBR.GDP, data = merge10)

M4 <- lm(NIK.Close.change ~ L.JPN.GDP, data = merge10)
```

```{r, echo=FALSE, results="asis"}
stargazer(M1, M2, M3, M4,
          type="html",
          title = "Basic Model",
          covariate.labels = c("France GDP growth", "Germany GDP growth", "Great Britain GDP growth", "Japan GDP growth"),
          dep.var.labels=c("CAC40","DAX", "FTSE100", "NIKKEI"),
          header = FALSE,
          no.space = TRUE)
```

